### 2018-06-12 express2018/equa.tex
text

We write $E[\til P]$ for the expression resulting from $E$ by
replacing each variable $X_i$   with the process $P_i$, assuming
$\til P$ and $\til X$ have the same length. (This is syntactic
replacement, akin to the substitution of the holes of a context with 
 processes.) 
The components of $\til P$ need not be
 different from each other, as it must be for the variables $\til X$.
If the system has infinitely many equations,
the  tuples $\til P$ and $\til X$
 are infinite too.
\begin{definition}
Suppose  $\{  X_i = E_i\}_{i\in I}$ is a system of equations: 
\begin{itemize}
\item
 $\til P$ is a \emph{solution of the 
system of equations  for $\wb$} 
if for each $i$ it holds
that $P_i \wb E_i [\til P]$.

@@

Examples of systems with a  unique solution for $\wb$ are: 
\begin{enumerate}
\item
$ X = a. X$

\item 
$ X_1 = a.  X_2$, $ X_2 = b.  X_1$

\end{enumerate}

The unique solution of the system (1), modulo $\wb$,  is the constant $K \Defi a
. K$:  for any other solution $P$ we have $P \wb K$.
The unique solution of (2), modulo $\wb$, are the constants $K_1 , K_2$
with $K_1 \Defi a . K_2$ and $K_2 \Defi b. K_1$; again, for any other
pair of solutions $P_1,P_2$ we have $K_1 \wb P_1$ and $K_2 \wb P_2$.
Examples of systems that do not have a unique solution are: 
\begin{enumerate}
\item $X = X $

\item $X = \tau . X$
\item $X = a | X$

\end{enumerate} 
All processes are solutions of (1) and (2); examples of solutions for
(3) are $K$ and $K | b$, for $K \Defi a
. K$.

@@

In other words,  
if the system is sequential, then 
 for
every expression $E_i$, any subexpression of $E_i$ in which $X_j $ 
appears, apart from $X_j$ itself,  is a sum (of prefixed terms). 
For instance, 
\begin{itemize}
\item $X = \tau. X + \mu . \nil$ is sequential but not 
 guarded, because the guarding prefix for the variable
is not visible.

\item $X =  \ell . X | P$ is  guarded but not sequential.

\item $X =  \ell . X + \tau. \res a (a .\outC b | a.\nil)$, as well as 
$X = \tau . (a. X + \tau . b .X + \tau  )$
are both 
 guarded and sequential.
\end{itemize}

@@

The proof exploits an invariance property on immediate transitions for
guarded and sequential expressions, and then extracts a bisimulation
(up-to bisimilarity) out
of the solutions of the system.  
To see the  need of the
 sequentiality  condition, 
  consider
 the equation (from \cite{Mil89}) 
\[ X = \res a (a. X | \outC a) \]
where $X$ is guarded but not sequential. Any processes that does
not use  $a$
 is a solution.

>>>>

We write $E[\til P]$ for the expression resulting from $E$ by
replacing each variable $X_i$   with the process $P_i$, assuming
$\til P$ and $\til X$ have the same length. (This is syntactic
replacement.) 
% The components of $\til P$ need not be
%  different from each other, as it must be for the variables $\til X$.
% If the system has infinitely many equations,
% the  tuples $\til P$ and $\til X$
%  are infinite too.
\begin{definition}
Suppose  $\{  X_i = E_i\}_{i\in I}$ is a system of equations: 
\begin{itemize}
\item
 $\til P$ is a \emph{solution of the 
system of equations  for $\wb$} 
if for each $i$ it holds
that $P_i \wb E_i [\til P]$.

@@

% Examples of systems with a  unique solution for $\wb$ are: 
% \begin{enumerate}
% \item
% $ X = a. X$

% \item 
% $ X_1 = a.  X_2$, $ X_2 = b.  X_1$

% \end{enumerate}

For instance, the solution of the equation 
$ X = a. X$ 
is  the process
$R \defi \recu A . a A$, and   for any other solution $P$ we have $P \wb R$.
In contrast, the equation 
 $X = a|  X$ has solutions that may be quite different, namely any process capable of
continuosly  performing $a$ action (but that could behave arbitrarily  on other actions).

% The unique solution of the system (1), modulo $\wb$,  is the constant $K \Defi a
% . K$:  for any other solution $P$ we have $P \wb K$.
% The unique solution of (2), modulo $\wb$, are the constants $K_1 , K_2$
% with $K_1 \Defi a . K_2$ and $K_2 \Defi b. K_1$; again, for any other
% pair of solutions $P_1,P_2$ we have $K_1 \wb P_1$ and $K_2 \wb P_2$.
% Examples of systems that do not have a unique solution are: 
% \begin{enumerate}
% \item $X = X $

% \item $X = \tau . X$
% \item $X = a | X$

% \end{enumerate} 
% All processes are solutions of (1) and (2); examples of solutions for
% (3) are $K$ and $K | b$, for $K \Defi a
% . K$.

@@

% In other words,  
% if the system is sequential, then 
%  for
% every expression $E_i$, any subexpression of $E_i$ in which $X_j $ 
% appears, apart from $X_j$ itself,  is a sum (of prefixed terms). 
% For instance, 
% \begin{itemize}
% \item $X = \tau. X + \mu . \nil$ is sequential but not 
%  guarded, because the guarding prefix for the variable
% is not visible.

% \item $X =  \ell . X | P$ is  guarded but not sequential.

% \item $X =  \ell . X + \tau. \res a (a .\outC b | a.\nil)$, as well as 
% $X = \tau . (a. X + \tau . b .X + \tau  )$
% are both 
%  guarded and sequential.
% \end{itemize}

@@

% The proof exploits an invariance property on immediate transitions for
% guarded and sequential expressions, and then extracts a bisimulation
% (up-to bisimilarity) out
% of the solutions of the system.  
To see the  need of the
 sequentiality  condition, 
  consider
 the equation (from \cite{Mil89}) 
\[ X = \res a (a. X | \outC a) \]
where $X$ is guarded but not sequential. Any processes that does
not use  $a$
 is a solution.

### 2018-06-12 express2018/expansion.tex
I have to use labels without hypen and underscore (don't know why) to prevent strange LaTeX compilation errors.
text

Bisimilarity is not congruence w.r.t direct sums, one obvious way to
fix the definition is to define another relation on top of it:
\begin{definition}{(rooted bisimilarity)}
Two processes $P$ and $Q$ are rooted bisimilar, denoted $P
\approx^c Q$, if and only if for all $\mu$ (including $\tau$):
\begin{enumerate}
 \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \wb Q'$;
\item the converse of (1) on the actions from $Q$.
\end{enumerate}
\end{definition}
Besides reducing the rooted bisimiarity of two processes to
the bisimilarities of their first-step transition ends, this definition also brings a proof technique for proving the
rooted bisimiarity by constructing a bisimulation:
\begin{lemma}{(Rooted bisimilarity by constructing a bisimulation)}
\label{l:obs_congr_by_weak_bisim}
Given a (weak) bisimulation $\RR$, if two processes $P$ and $Q$
satisfies the following properties:
\begin{enumerate}
 \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \RR Q'$;
\item the converse of (1) on the actions from $Q$.
\end{enumerate}
then $P$ and $Q$ are rooted bisimilar, i.e., $P \approx^c Q$.
\end{lemma}

>>>>

Bisimilarity is not congruence w.r.t direct sums, one obvious way to
fix the definition is to define another relation on top of it:
\begin{definition}[rooted bisimilarity]
Two processes $P$ and $Q$ are \emph{rooted bisimilar}, written $P
\approx^c Q$, if we have:
\begin{enumerate}
 \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
   \Arr\mu Q'$ and $P' \wb Q'$;
\item the converse of (1) on the actions from $Q$.
\end{enumerate}
\end{definition}
% Besides reducing the rooted bisimiarity of two processes to
% the bisimilarities of their first-step transition ends, this definition also brings a proof technique for proving the
% rooted bisimiarity by constructing a bisimulation:
% \begin{lemma}{(Rooted bisimilarity by constructing a bisimulation)}
% \label{l:obsCongrByWeakBisim}
% Given a (weak) bisimulation $\RR$, if two processes $P$ and $Q$
% satisfies the following properties:
% \begin{enumerate}
%  \item  $P \arr\mu P'$ implies that there is $Q'$ such that $Q
%    \Arr\mu Q'$ and $P' \RR Q'$;
% \item the converse of (1) on the actions from $Q$.
% \end{enumerate}
% then $P$ and $Q$ are rooted bisimilar, i.e., $P \approx^c Q$.
% \end{lemma}

### 2018-06-12 express2018/unique.tex
text

\begin{definition}
\label{d:guarded}
A system of contractions $\{  X_i \mcontr E_i\}_{i\in I}$
 is
\emph{weakly guarded}
if,  in each    $E_i$, each occurrence of
a \behavC\ variable is underneath a prefix.
\qed\end{definition}

\begin{lemma}
\label{l:uptocon}
Suppose $\til P$ and $\til Q$ are solutions  for $\wb$
 of a system of weakly-guarded
contractions.
For any context $\qct$, 
if  $\ct{\til P}\Arr{\mu}  R$,
 then 
there is a  context $\qctp$
such that $R \mcontrBIS \ctp{\til P}$ and  $\ct{\til Q} \Arcap{\mu}
 \wb \ctp{\til Q}$.
\end{lemma}

\finish{sopra: sum-free}

>>>>

\begin{definition}
\label{d:guarded}
A system of contractions $\{  X_i \mcontr E_i\}_{i\in I}$
 is
\emph{weakly guarded}
if,  in each    $E_i$, each occurrence of
a \behavC\ variable is underneath a prefix.

The system use \emph{weakly-guarded sums} if 
each $E_i$ only makes use of guarded sums.
\qed\end{definition}

\finish{introduce guarded sums, and contrast it with arbitrary sums}

\begin{lemma}
\label{l:uptocon}
Suppose $\til P$ and $\til Q$ are solutions  for $\wb$
 of a system of weakly-guarded
contractions that uses  weakly-guarded sums.
For any context $\qct$, 
if  $\ct{\til P}\Arr{\mu}  R$,
 then 
there is a  context $\qctp$
such that $R \mcontrBIS \ctp{\til P}$ and  $\ct{\til Q} \Arcap{\mu}
 \wb \ctp{\til Q}$.
\end{lemma}

\finish{here a sketch of the proof (notably with the induction argument)}

\begin{theorem}[unique solution of contractions for $\wb$]
\label{t:contraBisimulationU}
A system of weakly-guarded contractions
 that uses  weakly-guarded sums    has 
a unique solution 
 for $\wb$.
\end{theorem}
